---
title       : A Simple Farm Model for Saskatchewan
subtitle    : Based on PMP and Simulation 
author      : Jon Duan
job         : ECON 595 Agriculture Economics Project
logo        : uvic_shield.png
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax]            # {mathjax, quiz, bootstrap}
url:
  lib: librariesNew
  assets: assets
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
ext_widgets : {rCharts: ["libraries/highcharts","libraries/nvd3", "libraries/morris", "libraries/leaflet", "libraries/rickshaw"]}
---

<style>
strong {
font-weight: bold;
}
em {
font-style: italic;
}
</style>
## Motivation: A Model is Essential for Policy Analysis

### Impact of policies
 
###     Farm's
<center>  **REVENUE** </center>

###     Prodcution
<center> **COST**</center>

###     Crop
<center> **YIELD**</center>

###     Crop
<center> **PRICE**</center>




---
## Assumption of the Model
1. **Cost** is **fixed**
 - Cost is based on the farmers' operation and the situation of farm or soil. For cerntain area and time period, we assume it fixed. 
 
2. **Yield** is **normally** distributed
 - Yiled depends on something like weather or other random factors. 
 
3. **Price** following a **stochastic** process.
 - Price changes frequently, even in short term, like other commodity price.


---  
## Data Source and Description

- Five major crops in SK:  **Wheat, Oats, Barley, Flax, Canola**  

- The **monthly price** data is from **Statistics Canada** [_"Estimated areas, yield, production, average farm price and total farm value of principal field crops, in imperial units, annually 001-0017"_](http://www5.statcan.gc.ca/cansim/a26?lang=eng&retrLang=eng&id=0010017&tabMode=dataTable&srchLan=-1&p1=-1&p2=35).

 * The data covers 1985-12-01 to 2014-03-01
 
- The **annual yield** data is from  **Statistics Canada** [_"Farm product prices, crops and livestock, monthly Description 	002-0043"_](http://www5.statcan.gc.ca/cansim/a05?lang=eng&id=0020043&pattern=0020043&searchTypeByValue=1&p2=35)

 * The data covers 1991 to 2013 
 
- The **cost** per acre data is from **Government of Saskatchewan** [_"Crop Planning Guide"_](http://www.agriculture.gov.sk.ca/crop-planning-guides)

 * The data covers different soil zones, rotations, estimated yield, variable cost, fix cost.  

- The **farm** data from **Statistics Canada** [_The 2011 Census of Agriculture_](http://www.statcan.gc.ca/pub/95-640-x/2012002/prov/47-eng.htm) 
 
 * The data covers Farm numbers, type, size, receipts, area.


---
## A Representative Farm in SK
- A farm in __black__ soil zone in SK.
    - Farm size: 1668 acre ( __The 2011 Census of Agriculture__);  
    - Average price over three years per bussels ( **Statistics Canada** );   
    - Estimated yield (__"Crop Planning Guide"__);
    - Variable cost (__"Crop Planning Guide"__).
    - Land use (**Statistics Canada**)

```{r table1, comment=NA, results='asis',echo=FALSE}
setwd("E:/Dropbox/book/economics/485/projects/nlfarm/finalcode/slides")
input<- read.csv("../NLinput.csv", header = TRUE, sep = ",")
# print table for input
library(xtable)
print(xtable(input, caption= "Based on SK data"), type = "html", include.rownames = FALSE)
```


---
## Soil Zones in SK

<img class=center src=assets/img/Soil_Zones_in_SK.jpg height=572>

---
## The Cost of Production: PMP Approach
- Calibration method: **Postive Mathematical Programming**
- Advantage: 
    * Minimal __data__ requirement
    * Calibrate MP models exactly to __observed behaviour__
    * Optimum: combination of __binding constraints__ and first-order conditions
    * Policy analysis: __prediction__ of consequences and __sensitivity__ analysis 

- Three stages ( formalized by Howitt (1995a) )
    * Estimate output __marginal cost__
    * Estimate __cost function__
    * Policy analysis
<<<<<<< HEAD
>>>>>>> d2eb8b8d1eb13e2944dead8cc1352b8c3fd3b586



---
<<<<<<< HEAD
## PMP


---
## Simlation
GBM(http://quant.stackexchange.com/questions/4589/how-to-simulate-stock-prices-with-a-geometric-brownian-motion)
$$dS = \mu S dt + \sigma Sdz$$

to Lemma : Loosely speaking, if a random variable x follows an Ito process : (drift = a(x,t) et variance = b(x,t)2):

$$dx = a(x,t) dt + b(x,t) dz$$
 Solving the SDE over a single interval will still allow a conditional formula such as:
$$S_t = S_{t-1} \exp \{ (\mu - \sigma^2/2)\Delta t + \sigma (W_t - W_{t-1})\}$$
Solution:
$$S(t) = S(0) \exp{(\mu - \frac{\sigma^{2}}{2})t + B_{t}}$$
$$S_t = a \exp\left[\left(\mu - \frac{\sigma^2}{2}\right) t + \sigma W_t \right]$$
where Bt is a brownian motion

---
## Conclusion




=======
=======



---
>>>>>>> 37757b9c69b63b2a9d4e6dd342f12b778686a990
## Estimate Output Marginal Cost
- Maximize farmer's __profit__ including a set of __calibration__ constraints.

$$Max: R = \sum_{k=1}^n ( p_k x_k y_k - c_k x_k)$$
$$Subjet \; to: \sum_{k=1}^n x_k \le 1668 \;\; (1)$$ 
$$x_k \ge 0$$
$$x_k \le x_k^{obs} + 0.01,\; \forall k; \; \; [\lambda_k] \;\; (2)$$   
    
- $p_k$ is price, $x_k$ is land use, $y_k$ is yield,  $c_k$ is cost.  
- Where $(1)$ is __nature resource__ constraint (land 1668),
- $(2)$ is the __calibration__ constraint. Solve the  problem in GAMS $\to$ the associated __shadow price__ $\lambda_k$ for each crop.  



---
## Estimate cost function
- Assumption: a __quadratic__ cost function: $c_k = a x_k + b x_k^2$  for SK.  
$$\hat b_k = 2 \times \frac{\lambda_k}{x_k^{obs}} \; \text{and} \; \hat a_k=c_k^{obs} - 0.5 \times \hat b_k \times x_k^{obs} $$

```{r table2, comment=NA, results='asis',echo=FALSE}
shadowad<- read.csv("../shadowad.csv", header = TRUE, sep = ",")
# print table for shadow price
library(xtable)
print(xtable(shadowad, caption= "Based on SK data"), type = "html", include.rownames = FALSE)
```

- Replace the $c_k x_k$ in the objective functioin $\to$ solve the __revised__ problem in GAMS $\to$ duplicate the observed results $\to$ cost functions for SK are __calibrated__ .





---
## Cost, Yield, and Price
- Based on these cost functions $a_k x_k - b_k  x_k^2$, if we want to evaluate the impact of policy,need to look at $p_k x_k y_k$
     - Historial information 
     - Monte Carlo simulation for the price $p_k$ and yield $y_k$. (Turvey, 2012).
          $$R_{ij} = P_{ij} Y_{ij} - C_i$$
          $$Y_{ij} \sim N(E[Y_i], \sigma(Y_i))$$
          $$P_{ij} = P_{i0} e^{((\mu - \frac{1}{2}) \frac{7}{12} + \sigma N(0,1) \sqrt{\frac{7}{12}})}$$
          
- where $C_i$ is the variable cost associated with each crop  
- $Y_{ij}$ is crop yield generated from a normal distribution  















---
## Crop Yield History Data in SK (5 Year Mean)


```{r table3, echo = FALSE, results = 'asis'}

# recent mean of yield 5 year
# Read yield data, firs two row are  comment, 
yield<-read.csv("../yield.csv", skip =2,  header = T, sep = ",")
# Just take 5 crops and delete 1990 one NA
yield <- yield[-1, c(1:5,7)]
# correct names
names(yield)[c(4,6)]<-c("Flax","Wheat")

#library(xtable)
yield<-yield[,c("date","Wheat","Oats","Barley","Flax","Canola")]

meanyield <- apply(yield[19:23,-1],2, mean)
meanyield <- t(as.data.frame(meanyield))
print(xtable(meanyield), type = "html", include.rownames = T)

```

```{r, echo = FALSE, results = 'hide',message=FALSE}
require(rCharts)
```


```{r graph2, comment=NA,echo=FALSE,results='asis',message=FALSE}

yld <- transform(yield, date = as.character(date))
m1 <- mPlot(x = "date", y = c("Wheat","Oats","Barley","Flax","Canola"), type = "Line", data = yld)
m1$set(pointSize = 1, lineWidth = 2)
m1$save('assets/img/m1.html', cdn = TRUE)
cat('<iframe src="assets/img/m1.html" width=100%, height=600></iframe>')

```








---
## Simulation Crop Yield for SK model
- Since yield data is a time series, in order to find the __$\sigma$__ of yield, we need to detrend the yield data (Coble, 2013).
 - The easist way to detrend:
 $$Y_i = \beta_0 + \beta_1 t_i + \beta_2 t_i^2+ \; ... \; +\epsilon_i$$
 - Run a regression of yield on time using a polynomial form 
 - Then the predict yield $\hat Y_i$ is time trend
 - The standard deviation __$\sigma$__ of residual __$\epsilon$__ is what we want
 - For example, on next page the left graph is time trend for Wheat yield; the right is the residual.
 




---
## Detrand: time trend and Residual

```{r graph3, comment=NA,echo=FALSE, fig.width=14,message=FALSE}
library(forecast)
source("decomp.r")
#
decomwheat<-decomp(yield[,2], FALSE)
# set 2 graphs
par(mfrow=c(1,2))

plot(1991:2013,decomwheat$trend, type="o", ylab= "Time trend",main="Time trend of Wheat yield")
plot(1991:2013,decomwheat$remainder, type="o", ylab= "Residual", main="Residual of Wheat yield")
# set back to 1 graph
par(mfrow=c(1,1))

```



---
## 1000 Yield Simulation Based on 5 Year Mean and Detrand Standard Deviation

```{r table5, echo = FALSE, results = 'asis'}
sdDtrend <- function(x) sd(decomp(x, FALSE)$remainder)
sdyld <- apply(yield[,-1],2, function(x) sdDtrend(x))

# recent mean of yield 5 year
meanyld <- apply(yield[19:23,-1],2, mean)

Detrendsd <- sdyld

Detrendsd <- t(as.data.frame(Detrendsd))
print(xtable(Detrendsd), type = "html", include.rownames = T)
```



```{r  graph5, comment=NA,echo=FALSE,message=FALSE,fig.width=14,fig.height=6}
library(ggplot2)

#simnum=1000
#crops=5        
#simYld <- array(0, dim=c(simnum, crops))

## yield random 
#for (i in 1:simnum) {
#        simYld[i,] <- rnorm(crops, meanyld, sdyld)
#}
#simYld<-as.data.frame(simYld)
#names(simYld)<-names(yield)[-1]

simYld<-read.csv("../simYld.csv")
ggplot(simYld, aes(x=1:nrow(simYld))) + geom_line(aes(y = simYld$Wheat, colour = "Wheat")) +  geom_line(aes(y = simYld$Barley, colour = "Barley"))+ geom_line(aes(y = simYld$Oats, colour = "Oats")) + geom_line(aes(y = simYld$Flax, colour = "Flax"))  + geom_line(aes(y = simYld$Canola, colour = "Canola"))+ labs(title="Simulative Yield", x= "State", y="Price") + theme(legend.title=element_blank())                                                                                                                                                                                                                               
                                                                                                                                                                                                                                                                                                                     
                                                                                                                                                                                                                                                                                                                      
```









---
## Crop Price History Data of SK


```{r table4, echo = FALSE, results = 'asis'}

price<- read.csv("../price.csv", header = T, sep = ",")

# set the date format
price[, 1] <- as.Date(price[, 1], format = "%d/%m/%Y")
# print table for price

meanprice<- apply(price[265:340,-1],2, mean)
meanprice <- t(as.data.frame(meanprice))
print(xtable(meanprice), type = "html", include.rownames = T)

```



```{r graph4, comment=NA,echo=FALSE,results='asis',message=FALSE}

pr <- transform(price, date = as.character(date))
m2 <- mPlot(x = "date", y = c("Wheat","Oats","Barley","Flax","Canola"), type = "Line", data = pr)
m2$set(pointSize = 1, lineWidth = 2)
m2$save('assets/img/m2.html', cdn = TRUE)
cat('<iframe src="assets/img/m2.html" width=100%, height=600></iframe>')

```













---
## Price Simulation: Geometric Brownian Model

- Recall
$$P_{ij} = P_{i0} e^{((\mu - \frac{1}{2}) \frac{7}{12} + \sigma N(0,1) \sqrt{\frac{7}{12}})}$$
                 
      - $P_{i0}$ is the initial spring price as of March 2014; 
      - Price generation is based on a 7-month growing season;
      - $P_{ij}$ is the random commodity harvest price generated by a log normal (Brownian) process with: 
       - drift $\mu$, 
       - volatility $\sigma$, 
       - random deviate drawn from a normal distribution with zero mean and variance of 1.0;

---
## Price Simulation: Geometric Brownian Model(Cont.)
- Geometric Brownian Model follow this equation:
$$dS_t = \mu S_t dt + \mu S_t dW_t$$

 - $\mu S_t dt$ is **deterministic part**;    $\mu S_t dW_t$ is **stochastic part**.

 - $dW_t$ is the Brownian motion, which follows random normal distributioin $N(0,t)$.

 - $\mu$ is **drift**; $\sigma$ is **diffusion**. $\sigma$ increases the amount of randomness entering the system.


```{r table6, comment=NA, echo=FALSE, results='asis'}
gbmPara<- read.csv("../gbmPara.csv")
print(xtable(gbmPara,caption="Estimated Mu and Sigma based on historical data in SK",digits=c(0,8,9)), type = "html", include.rownames = TRUE)

````















---
## 1000 Crop Price Simulation for SK model

```{r graph6, comment=NA,echo=FALSE, fig.width=14,message=FALSE}

simPr<-read.csv("../simPr.csv")

ggplot(simPr, aes(x=1:nrow(simPr))) + geom_line(aes(y = simPr$Wheat, colour = "Wheat")) +  geom_line(aes(y = simPr$Barley, colour = "Barley"))+ geom_line(aes(y = simPr$Oats, colour = "Oats")) + geom_line(aes(y = simPr$Flax, colour = "Flax")) + geom_line(aes(y = simPr$Canola, colour = "Canola"))+ labs(title="Simulative Price", x= "State", y="Price")+ theme(legend.title=element_blank())

```














---
## Conclusion
- With the cost, yield, and price data, the SK farm model can be structured as follows:

$$Min: \sigma_p^2 = \frac{1}{m} \sum_{j=1}^m ( \pi_j - E[\pi] )^2$$
$$Subjet \; to: \sum_{i=1}^n x_k = 1668 \; $$ 
$$\sum_{i=1}^n E[R_i] x_i = K \; (2)$$

- $\sigma_p^2$ measures the unstability of farmers' income ; $\pi$ is the profit for one state
- (2) $K$ is the target profit
- $E[R]$ is expect the average profit per acre
- $m$ is 1000 simulation state



---
## Conclusion
- The impact of policy can be added to model. For example,whole farm insurance:

$$\sum_{i=1}^n R_{1,i} x_i + Max[Z - \sum_{i=1}^n R_{1,i} x_i, \; 0]- \frac{\delta}{m} \sum_{j=1}^m Max[Z - \sum_{i=1}^n R_{j,i} x_i, \; 0] - \pi_1 = 0 $$
$$\sum_{i=1}^n R_{m,i} x_i + Max[Z - \sum_{i=1}^n R_{m,i} x_i, \; 0]- \frac{\delta}{m} \sum_{j=1}^m Max[Z - \sum_{i=1}^n R_{j,i} x_i, \; 0] - \pi_m = 0 $$

- $Z$ is the income coverage level to be protected by insurance such as 70% farmers history income
- $Max[Z - \sum_{i=1}^n R_{1,i} x_i, \; 0]$ is indemnity payout that farmer can get from insurance
- $\frac{\sigma}{m} \sum_{j=1}^m Max[Z - \sum_{i=1}^n R_{j,i} x_i, \; 0]$ is permium that farmers need to pay.
- $\frac{\sigma}{m}$ is subsidy rate. If $\delta$ = 0.50, the premium is subsidized by 50 percent

   

