SK Model
========================================================

## data
I got the monthly price data of crops in Alberta from Statistics Canada, [Farm product prices, crops and livestock, monthly         Description (dollars per metric tonne) 002-0043](http://www5.statcan.gc.ca/cansim/a05?lang=eng&id=0020043&pattern=0020043&searchTypeByValue=1&p2=35).



```{r table1, comment=NA, results='asis'}
setwd("E:\\Dropbox\\Github\\sk")
price<- read.csv("price.csv", header = T, sep = ",")

# set the date format
price[, 1] <- as.Date(price[, 1], format = "%d/%m/%Y")
# print table for price
library(xtable)
print(xtable(summary(price)), type = "html", include.rownames = F)

```


```{r ,echo =FALSE, message=FALSE, results = 'hide'}
library(reshape)
#library(plyr)
#with(melt(price[,-1]), tapply(value, factor(variable), mean))
# get mean and standard deviation for PMP from recent 6 year
#str(price)
meanpr<-apply(price[c(265:340),-1], 2, mean)
sdpr<-apply(price[c(265:340),-1],2, sd)
#summary(price)
```
Size of Farms The 2011 Census of Agriculture recorded 36 952 Saskatchewan census farms as of May 10, 2011 a decline of 16.6 per cent from 44 329 census farms as at May 15, 2006.
The total farm land area was recorded at 61.6 million acres – averaging 1 668 acres per farm. In 2006, the average area was 1 449 acres. On average, Saskatchewan has the largest farms in Canada.

### Plot of price
```{r}

#install.packages("reshape","ggplot2")
#boxplot(price[c(300:340),-1])

library(ggplot2)
histpr<-ggplot(data=melt(price[,-1]), aes(as.factor(variable), value, fill=factor(variable)))
histpr + geom_boxplot() + guides(fill=guide_legend(title=NULL))+labs(title="Boxplot for History Price", x= "Crop", y="Price")
# line

ggplot(price, aes(price[,1])) + geom_line(aes(y = price[,2], colour = "Wheat")) +  geom_line(aes(y = price[,3], colour = "Barley")) + geom_line(aes(y = price[,4], colour = "Oats")) + geom_line(aes(y = price[,5], colour = "Flax")) + geom_line(aes(y = price[,6], colour = "Calona")) + labs(title="Trend of History Price", x= "Date", y="Price")+ theme(legend.title=element_blank())

```

### Get the yield data.

```{r table2, comment=NA, results='asis'}
# Read yield data, firs two row are  comment, 
yield<-read.csv("yield.csv", skip =2,  header = T, sep = ",")

# Just take 5 crops and delete 1990 one NA
yield <- yield[-1, c(1:5,7)]
# correct names
names(yield)[c(4,6)]<-c("Flax","Wheat")
#names(yield)
#yield$date<-as.Date(yield$date)
#library(xtable)
yield<-yield[,c("date","Wheat","Oats","Barley","Flax","Canola")]
print(xtable(summary(yield)), type = "html", include.rownames = FALSE)
#summary(yield)
head(yield)

```

## Plot of yield
```{r}
# boxplot(yield[,-1])
library(reshape)
library(ggplot2)
#library(scales)
histyldb<-ggplot(data=melt(yield[,-1]), aes(as.factor(variable), value, fill=factor(variable)))
histyldb + geom_boxplot() + guides(fill=guide_legend(title=NULL))+labs(title="Boxplot for History Yield", x= "Crop", y="yield")
# line
histyldl <- ggplot(yield, aes(yield[,1]))
yldBarley <- geom_line(aes(y = yield[,2], colour = "Barley"))
yldCanola <- geom_line(aes(y = yield[,3], colour = "Canola"))
yldFlax<- geom_line(aes(y = yield[,4], colour = "Flax"))
yldOats<- geom_line(aes(y = yield[,5], colour = "Oats"))
yldWheat<-geom_line(aes(y = yield[,6], colour = "Wheat"))

histyldl + yldBarley + yldCanola + yldFlax + yldOats+ yldWheat + labs(title="Trend of History yield", x= "Date", y="yield")+ theme(legend.title=element_blank())

```



## Detrand yield and get varibility (standard deviation) of yield for simulation

```{r ,message =FALSE}
library(forecast)
source("decomp.r")
str(yield)
sdDtrend <- function(x) sd(decomp(x, FALSE)$remainder)

#sd(decomp(yield$Wheat)$remainder)
#sd(decomp(yield[,3], FALSE)$remainder)
   
sdyld0 <- apply(yield[,-1],2, function(x) sd(x))
sdyld <- apply(yield[,-1],2, function(x) sdDtrend(x))

# recent mean of yield 5 year
meanyld <- apply(yield[19:23,-1],2, mean)
#decomyld<-decomp(yield[,2], FALSE)
#plot(1991:2013,decomyld$trend, type="o")
#plot(1991:2013,decomyld$remainder, type="o" )

```



## yield simulation


```{r ,message =FALSE}


```




## Yield Simulation


```{r ,results='asis'}
simnum=1000
crops=5        
simYld <- array(0, dim=c(simnum, crops))


# yield random 
for (i in 1:simnum) {
        simYld[i,] <- rnorm(crops, meanyld, sdyld)
}
simYld<-as.data.frame(simYld)
names(simYld)<-names(yield)[-1]
write.table(simYld,"simYld.csv", col.names = TRUE,row.names = FALSE, sep=",", )

```


## PMP

(ALPH(k)+0.5*BETA(k)*NX(k)))*NX(k))


## Price simulation 


### parameter estimate


#### GBM parameter estimate for price motion

```{r}
library(yuima)
# set gbm model
set.seed(12345)
        
ymodel <- setModel(drift=c("theta2*x"), diffusion=c("theta1*x"),time.variable="t", state.variable="x", solve.variable="x")
        
# initial gbm parament matrix        
gbmPara<- matrix(ncol=2)

# loop for all parameter
for(i in 2:ncol(price)){        
        x <- price[,i]
        Data2=setYuima(data=setData(x),ymodel)
        mle1 <- qmle(Data2, start = list(theta1 = 1, theta2 = -0.001),lower = list(theta1=-2, theta2=-1),upper = list(theta1=0.5, theta2=0.5), method = "L-BFGS-B")
        coef(mle1)
        gbmPara <- rbind(gbmPara, coef(mle1))
        }        
```


### Generate all parameter for all crop price

```{r table3, comment=NA, results='asis'}
# set gbmPara
gbmPara<-as.data.frame(gbmPara)
gbmPara<-gbmPara[-1,]
names(gbmPara)<-c("sigma","mu")
row.names(gbmPara) <- names(price)[-1]
# print GBM parameters talbe
print(xtable(gbmPara), type = "html", include.rownames = FALSE)
# set mean of price as initial value
# meanPr<-apply(price[265:340,-1],2,mean)
# print mean of price talbe
print(xtable(meanPr), type = "html", include.rownames = FALSE)
# write to talbe csv

write.table(gbmPara, "gbmPara.csv", col.names = TRUE,row.names = T, sep="," )
```


## simulation of crop price


```{r}

# set sampling, model, frequence
n <- 1000
ysamp <- setSampling(Terminal=(n)^(1/2), n=n) 
yuima <- setYuima(model=ymodel, sampling=ysamp)

# set simulative price matrix 1000x
simPr<-matrix(nrow=n+1)
# set initial price as March 2014
initPr<-price[340,-1]
#meanpr<-apply(price[c(265:340),-1], 2, mean)
# set loop for 5 crop sim price
for(j in 1:length(initPr)){
        dyuima <- simulate(yuima, xinit= as.numeric(initPr[j]), true.parameter=list(theta1=gbmPara[j,1],theta2=gbmPara[j,2]))
        as.numeric(dyuima@data@ original.data)
        simPr<-cbind(simPr,as.numeric(dyuima@data@ original.data))
}

```







### Generate simulation price matrix 1000*5


```{r table4, comment=NA, results='asis'}
# set gsimPr data.frame
#head(simPr)
# delete the NA in first column and first row  which is  s0
simPr<-as.data.frame(simPr[-1,-1])
names(simPr)<-names(price)[-1]

# print GBM parameters talbe
print(xtable(head(simPr)), type = "html", include.rownames = FALSE )
# set mean of price as initial value

write.table(simPr,"simPr.csv", col.names = TRUE,row.names = FALSE, sep=",", )

```


### Plot of simulative price
```{r}
# library(lattice)
#install.packages("reshape","ggplot2")
#boxplot(price[c(300:340),-1])
#library(reshape)
library(ggplot2)
simPrPlot<-ggplot(data=melt(simPr), aes(as.factor(variable), value, fill=factor(variable)))
simPrPlot + geom_boxplot() + guides(fill=guide_legend(title=NULL))+labs(title="Boxplot for Simulative Price", x= "Crop", y="Price")
# line

plot.ts(simPr)
plot.ts(price[,-1])
reset_index(simPr)
head(simPr)
ggplot(simPr, aes(x=1:nrow(simPr))) + geom_line(aes(y = simPr$Wheat, colour = "Wheat")) +  geom_line(aes(y = simPr$Barley, colour = "Barley"))+ geom_line(aes(y = simPr$Oats, colour = "Oats")) + geom_line(aes(y = simPr$Flax, colour = "Flax")) + geom_line(aes(y = simPr$Canola, colour = "Canola"))+ labs(title="Trend of Simulative Price", x= "Date", y="Price")+ theme(legend.title=element_blank())

```

## simulation revenue

```{r}


# clean price matrix
head(simPr)
str(simPr)
head(simYld)
# swap column to match price
simYld<-simYld[,c("Wheat","Oats","Barley","Flax","Canola")]
# define the revenue
simRev<-simPr[1,]*simYld[1,]
# generate revenue
for (i in 2:simnum) {        
	simRev <- rbind(simRev,(simPr[i,]*simYld[i,])) 
}

write.table(simRev,"simRev.csv", col.names = TRUE,row.names = FALSE, sep=",", )

#head(simRev)
s#im = sapply(1:simnum, function(x) simPr[x,]*simYld[x,])
#str(sim)
```


